---
title: "Hw 5"
author: "Andrew Maloney"
date: "4/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Part 1

The code below downloads the data for Ames only and saves a local copy to your machine. The first time you run this code, make sure you have a good internet connection and enough space on your machine (~50 MB). If the local copy exists, re-knitting the file will load the (binary) file from disk and be much faster.

```{r}
if(file.exists("alcohol.rds")) {
  data <- readRDS("alcohol.rds")
} else {
  url <- "https://xdaiisu.github.io/ds202materials/data/ames.csv"
  data <- readr::read_csv(url)
  saveRDS(data, file = "alcohol.rds")
  
}

```

```{r}

library(dplyr)
library(tidyverse)
library(stringr)

```

# Part 2 Data Cleaning

1. Extract geographic latitude and longitude from the variable Store Location
```{r}

data_clean_1 <- data %>% separate(`Store Location`, c("address1", "address2", "coords"), sep = "\n") %>% select(-address1, -address2) %>% mutate(coords = gsub("[()]", "", coords)) %>% separate(coords, c("lat", "long"), sep = ",") # My code

str(data_clean_1)

#Code using regular expressions
#data_clean_1 <- data %>% mutate(`Store Location` = gsub("(?<=\\()[^()]*(?=\\))(*SKIP)(*F)|.", "", `Store Location`, perl=T)) %>% separate(`Store Location`, c("x-coordinate", "y-coordinate"), sep = ",") 

```

2. Check variable types. Pick five variables that need to be converted to a different type and fix those.

3. Extract year, month and day from the variable Date

```{r}

library(lubridate)
table(data_clean_1$Pack)


data_clean_2 <- data_clean_1 %>% mutate(long = as.numeric(long), lat = as.numeric(lat), `Vendor Number` = as.numeric(`Vendor Number`), Pack = factor(Pack), `Bottle Volume (ml)` = factor(`Bottle Volume (ml)`))


#data_clean_x <- data_clean_2 %>% mutate(Date = mdy(Date))

data_clean_3 <- data_clean_2 %>% mutate(Date_New = mdy(Date)) %>% separate(Date, c("Month", "Day", "Year"), sep = "/")

```


# Part 3 First Overview:

1. Plot a scatterplot of lat and long of store locations.
```{r}
library(ggplot2)
library(plotly)

data_clean_3 %>% ggplot(aes(x = long, y = lat))+
  geom_point(na.rm = TRUE) + geom_jitter(na.rm = TRUE)+
  ggtitle("Scatterplot of Latitude Vs Longitude")+ xlab("Longitude")+ ylab("Latitude") 

```

2. Provide a visual breakdown of the liquor category (by Category Name). Include volume sold in the breakdown.
```{r}

data_clean_x <-  data_clean_3 %>% group_by(`Category Name`) %>% summarise(total_amount_of_alcohal_sold = (sum(`Volume Sold (Gallons)`))) %>% filter(total_amount_of_alcohal_sold >= 20000)

#summary(data_clean_x$total_amount_of_alcohal_sold)

data_clean_x %>% ggplot(aes(x = reorder(`Category Name`, total_amount_of_alcohal_sold), y = total_amount_of_alcohal_sold))+ geom_point()+ coord_flip()


```


# Part 4: 
Find the daily sales liquor sales in Ames in 2018. Summarize number of sales, volume of liquor sold and amount of money spent. Plot volume sold by day (use a scatterplot of volume by day and facet by month). Describe any patterns. Find the dates for ISU football home games in Fall 2018. Can you see a pattern?
```{r}
year_2018 <- data_clean_3 %>% mutate(Day = as.numeric(Day), Year = as.numeric(Year), Month = as.numeric(Month)) %>% filter(Year == 2018)

History_2018 <- year_2018 %>% group_by(Day) %>% summarize(Total_bottles_Sold = sum(`Bottles Sold`), Total_Money_Spent = sum(`Sale (Dollars)`), Total_Volume_Liq = sum(`Volume Sold (Liters)`))

History_2018 %>% ggplot(aes(x = Day, y = Total_bottles_Sold))+
  geom_point()+geom_line() 

History_2018 %>% ggplot(aes(x = Day, y = Total_Money_Spent))+
  geom_point()+geom_line()

History_2018 %>% ggplot(aes(x = Day, y = Total_Volume_Liq))+
  geom_point()+geom_line()

year_2018 %>% ggplot(aes(x = Day, y = `Volume Sold (Liters)`))+
  geom_point()+ geom_jitter()+ facet_wrap(~Month)


#year_2018football <- year_2018 %>% filter(Month == 9 & Day == 8) & (Month == 9 & Day == 15) & (Month == 9 & Day == 22) & (Month == 9 & Day == 29) & (Month = 10 & Day == 6) & (Month == 10 & Day == 13) & (Month = 10 & Day == 27) & (Month == 11 & Day == 3) & (Month == 11 & Day == 10) & (Month == 11 & Day == 17) & (Month == 11 & Day == 24) & (Month == 12 & Day == 1) & (Month == 12 & Day == 28)) 



```